---
title: "Data wrangling 2"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(rvest)
library(httr)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "virids"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## strings and vectors 

```{r}
string_vec = c("my", "name", "is", "jeff")

str_detect(string_vec, "m")
str_detect(string_vec, "jeff")
str_detect(string_vec, "Jeff")

str_replace(string_vec, "jeff", "Jeff")
str_replace(string_vec, "e", "This is an uppercase E")
str_replace(string_vec, "jeff", "") 

```

^ represents beginning of the line
  - finding "i think" at the beginning of the line
$ represents end of the line

````{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "i think")
str_detect(string_vec, "^i think")
str_detect(string_vec, "i think$")

```

[]to inlcude uppercase and lowercase (enter both into brackets)

```{r}
string_vec = c(
  "Y'all remember Pres. HW Bush?",
  "I saw a green bush",
  "BBQ and Bushwalking at Molonglo Gorge",
  "BUSH -- LIVE IN CONCERT!!"
  )

str_detect(string_vec, "Bush")
str_detect(string_vec, "[Bb]ush")

```

specify a range 

```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "[0-9][a-zA-Z]")
```

"." to represent anything 

```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11")
str_detect(string_vec, "711")
str_detect(string_vec, "7.")
```


specify for "[ or \"
```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, "\\[")
```


## Why factors are weird 

* without specifcying, factor will level in numeric order

* convert back to numeric, male will be "2" and female "1" 

* fct_relevel: reorder, just state one = that will be moved up to the beginning 

```{r}
factor_vec = factor(c("male", "male", "female", "female"))

as.numeric(factor_vec)

factor_vec = fct_relevel(factor_vec, "male")
as.numeric(factor_vec)
```

## NSQH

* downlaod url 

* select first table only and remove first row

* convert values to numeric (converted to character because of presence of "a")



```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

table_marj = 
  read_html(nsduh_url) %>% 
  html_table() %>% 
  first() %>% 
  slice(-1)
```

Let's clean this up

* removing P value column 

* pivot longer

* separating age and year (telling to separate on "(")

* removing all parenthesis 

* removing letters at end of percent 

```{r}
marj_df = 
  table_marj %>% 
  select(-contains("P value")) %>% 
  pivot_longer(
    -State,
    names_to = "age_year",
    values_to = "percent"
    ) %>% 
  separate(age_year, into = c("age", "year"), "\\(") %>% mutate(
    year = str_replace(year, "\\)", ""), 
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)
  ) %>% 
filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
```

Do dataframe stuff

* states are in alphabetical order because factored in alphabetical order 
  - change in mutate to ordered in percent order
```{r}
marj_df %>% 
  filter(age == "12-17") %>% 
  mutate(
    State = fct_reorder(State, percent)
  ) %>% 
  ggplot(aes(x = State, y = percent, color = year)) + 
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
  
```


## Restaraunt inspections 

```{r}
data("rest_inspec")
```

```{r}
rest_inspec %>% 
  janitor::tabyl(boro, grade)
```

```{r}
rest_inspec = 
  rest_inspec %>% 
  filter(
   str_detect(grade, "[ABC]"),
   !(boro == "Missing")
  ) %>% 
  mutate(boro = str_to_title(boro))
```

```{r}
rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]"))%>% 
  janitor::tabyl(boro, grade)
```

```{r}
rest_inspec%>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]"))%>%
  mutate(
    boro = fct_infreq(boro)
  ) %>% 
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar()
```

Changing a factor label 

```{r}
rest_inspec%>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]"))%>%
  mutate(
    boro = fct_infreq(boro),
    boro = fct_recode(boro,"The City" = "Manhattan")
  ) %>% 
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar()

fct_

```

